---
title: "Hypothesis Testing for Credit Scoring Dataset"
author: "Gleb Sokolov"
date: '2022-05-06'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
library(PropCIs)
library(zeallot)
library(DBI)
library(vcd)
con <- DBI::dbConnect(odbc::odbc(), driver = "/usr/local/Cellar/psqlodbc/13.02.0000/lib/psqlodbcw.so", database = "yukontaf", UID = "glebsokolov", host = "localhost",
  port = 5432)
```

```{r}
credit_score <- dbSendQuery(con, "SELECT * FROM credit_score")
credit_score <- dbFetch(credit_score)
```


```{r}
for (col in 1:ncol(credit_score)) {
  colnames(credit_score)[col] <- tolower(colnames(credit_score)[col])
}
```

```{r}
credit_score <- subset(credit_score, select = -c(index, id))
categories <- c("sex", "education", "marriage", "default")
for (col in categories) {
  credit_score[, col] <- as.factor(credit_score[, col])
}
numerical <- names(subset(credit_score, select = -c(sex, education, marriage, default)))
for (n in numerical) {
  credit_score[, n] <- as.numeric(unlist(credit_score[, n]))
}
head(credit_score, 10)
```

First, I have to admit that this is a synthetical dataset, which means that there are no missing values, outliers, errors or any other mistakes, so these examinations will be skipped.

```{r}
credit_score$limit_bal <- log(credit_score$limit_bal)
p1 <- ggplot(credit_score, aes(x = limit_bal)) + geom_histogram()
ggplotly(p1)
```

```{r}
credit_score$age <- log(credit_score$age)
p2 <- ggplot(credit_score, aes(x = age)) + geom_histogram()
ggplotly(p2)
```


```{r}
p3 <- ggplot(credit_score, aes(x=sex, y=limit_bal)) + geom_boxplot()
ggplotly(p3)
```
```{r}
p4 <- ggplot(credit_score, aes(x=education, y=limit_bal)) + geom_boxplot()
ggplotly(p4)
```

```{r}
p5 <- ggplot(credit_score, aes(x=marriage, y=limit_bal)) + geom_boxplot()
ggplotly(p5)
```
```{r}
t <- (table(credit_score$default,credit_score$sex))
t <- as.data.frame(t)
colnames(t) <- c('default', 'sex', 'cnt')
p6 <-
  ggplot(t, aes(x = default, y = cnt, fill = sex)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p6)
```



```{r}
t <- (table(credit_score$default, credit_score$education))
t <- as.data.frame(t)
colnames(t) <- c('default', 'education', 'cnt')

p6 <-
  ggplot(t, aes(x = default, y = cnt, fill=education)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p6)

```

```{r}
t <- (table(credit_score$default, credit_score$marriage))
t <- as.data.frame(t)
colnames(t) <- c('default', 'marriage', 'cnt')

p7 <-
  ggplot(t, aes(x = default, y = cnt, fill=marriage)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p7)
```
Let's test two hypothesis: 
- Are the mean credit limits (limit_bal) value for two groups default = 0 (didn't returned the credit) and default = 1 equal to each other?
- Are the distributions of the limit_bal for these two groups also equal to each other?

In order to answer these and the following questions I will calculate **confidence intervals**.

```{r}
t.test(limit_bal ~ default, data = credit_score)
wilcox.test(limit_bal ~ default, data = credit_score)
```

**These results are obviously practically significant.**

Now, lets test another pair of hypothesis:
- Are the mean ages and their distributions for these two groups equal to each other?

```{r}
t.test(age ~ default, credit_score)
wilcox.test(age ~ default, credit_score)
```

The result we received tells us that statistically, mean ages *are* different, but from the confidence interval value we can see that **this difference is hardly practically signigicant.**

Now let's see if the gender composition for the two groups differ.

```{r}
good <- filter(credit_score, default == 0)
bad <- filter(credit_score, default == 1)
c(ngoodmen, total_good, nbadmen, total_bad) %<-% c(table(good$sex)[1], sum(table(good$sex)), table(bad$sex)[1], sum(table(bad$sex)))
diffscoreci(ngoodmen, total_good, nbadmen, total_bad, conf.level = 0.95)
```

That means that men do not return their credits **slightly more often** (3-6%) than women.

Now, let's see if the education level impacts default rate. First, calculate table which will show us the sizes of default and no-default groups for each education level, secondly, let's see how do these sizes differ from the expected ones, next calculate the value of the statistical criteria.

```{r}
crosstab <- table(credit_score$education, credit_score$default)
crosstab
crosstab - chisq.test(crosstab)$expected
chisq.test(crosstab)
assocstats(crosstab)
```

Finally, let's see if the marriage category impacts the default category.

```{r}
marriage_crosstab <- table(credit_score$marriage, credit_score$default)
marriage_crosstab
assocstats(marriage_crosstab)
```

For both variables (education and marriage) we see that **they statitically significant impact the default category**. However, the contigency coefficients (which tells us how strong the features are correlated) are relatively small.