---
title: "Hypothesis Testing for Credit Scoring Dataset"
author: "Gleb Sokolov"
date: '2022-05-06'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
library(PropCIs)
library(zeallot)
library(DBI)
library(vcd)
library(perm)
library(EnvStats)
library(lmPerm)
library(multcomp)
library(plyr)
con <- DBI::dbConnect(odbc::odbc(), driver = "/usr/local/Cellar/psqlodbc/13.02.0000/lib/psqlodbcw.so", database = "yukontaf", UID = "glebsokolov", host = "localhost",
  port = 5432)
```

```{r permutation test}
permutation.test <- function(treatment, outcome, n){
  original <- diff(tapply(outcome, treatment, mean))
  original <- original[complete.cases(original)]
  distribution=c()
  result=0
  for(i in 1:n){
    dist <- diff(by(outcome, treatment[sample(length(treatment), replace=FALSE)], mean))
    distribution[i] <- dist[complete.cases(dist)]
  }
  # result=sum(abs(distribution) >= abs(original))/(n)
  return(distribution)
}
```

```{r load data}
credit_score <- dbSendQuery(con, "SELECT * FROM credit_score")
credit_score <<- as.data.frame(dbFetch(credit_score))
```

```{r transform colnames}
for (col in 1:ncol(credit_score)) {
  colnames(credit_score)[col] <- tolower(colnames(credit_score)[col])
}
```

```{r make categorical cols}
credit_score <- subset(credit_score, select = -c(index, id))
categories <- c("sex", "education", "marriage", "default")
for (col in categories) {
  credit_score[, col] <- as.factor(credit_score[, col])
}
numerical <- names(subset(credit_score, select = -c(sex, education, marriage, default)))
for (n in numerical) {
  credit_score[, n] <- as.numeric(unlist(credit_score[, n]))
}
```

```{r}
credit_score$education = factor(credit_score$education, levels=c(6, 5, 4, 3, 2, 1, 0), ordered=TRUE)
```

```{r}
credit_score <- na.omit(credit_score)
```

First, I have to admit that this is a synthetical dataset, which means that there are no missing values, outliers, errors or any other mistakes, so these examinations will be skipped.

```{r limit_bal hist, include=FALSE}
credit_score$limit_bal_log <- log(credit_score$limit_bal)
p1 <- ggplot(credit_score, aes(x = limit_bal)) + geom_histogram()
hist1 <-
  permutation.test(credit_score$limit_bal,
                   as.numeric(credit_score$default),
                   1000)
hist <- ggplot(as.data.frame(hist1), aes(x = hist1)) + geom_histogram()
```

```{r}
ggplotly(p1)
```

Note: distribution of limit_bal is probably not normal. We will use this fact later.

```{r age hist}
credit_score$age <- log(credit_score$age)
p2 <- ggplot(credit_score, aes(x = age)) + geom_histogram()
ggplotly(p2)
```

```{r limit_bal~sex}
p3 <- ggplot(credit_score, aes(x=sex, y=limit_bal)) + geom_boxplot()
ggplotly(p3)
```

```{r limit_bal~education}
p4 <- ggplot(credit_score, aes(x=education, y=limit_bal, fill=education)) + geom_boxplot()
ggplotly(p4)
```

```{r limit_bal~marriage}
p5 <- ggplot(credit_score, aes(x=marriage, y=limit_bal, fill=marriage)) + geom_boxplot()
ggplotly(p5)
```

```{r default~sex rates}
t <- (table(credit_score$default,credit_score$sex))
t <- as.data.frame(t)
colnames(t) <- c('default', 'sex', 'cnt')
p6 <-
  ggplot(t, aes(x = default, y = cnt, fill = sex)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p6)
```

```{r default~education rates}
t <- (table(credit_score$default, credit_score$education))
t <- as.data.frame(t)
colnames(t) <- c('default', 'education', 'cnt')

p6 <-
  ggplot(t, aes(x = default, y = cnt, fill=education)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p6)

```

```{r default~marriage rate}
t <- (table(credit_score$default, credit_score$marriage))
t <- as.data.frame(t)
colnames(t) <- c('default', 'marriage', 'cnt')

p7 <-
  ggplot(t, aes(x = default, y = cnt, fill=marriage)) + geom_bar(stat = 'identity',  position = position_dodge())
ggplotly(p7)
```

Let's test two hypothesis: - Are the mean credit limits (limit_bal) value for two groups default = 0 (didn't returned the credit) and default = 1 equal to each other? - Are the distributions of the limit_bal for these two groups also equal to each other?

In order to answer these and the following questions I will calculate **confidence intervals**. Here and further I will use permutation non-parametric test in order to compare distributions

```{r}
t.test(limit_bal ~ default, data = credit_score)
wilcox.test(limit_bal ~ default, data = credit_score)
```

```{r}
hist1 = permutation.test(credit_score$default, credit_score$limit_bal, 1000)
p8 <- ggplot(as.data.frame(hist1), aes(x=hist1)) + geom_histogram()
ggplotly(p8)
```

**These results are obviously practically significant.**

Now, lets test another pair of hypothesis: - Are the mean ages and their distributions for these two groups equal to each other?

```{r}
t.test(age ~ default, credit_score)
wilcox.test(age ~ default, credit_score)
```

```{r age permutation test}
hist2 <- permutation.test(credit_score$default, credit_score$age, 1000)
p9 <- ggplot(as.data.frame(hist2), aes(x=hist2)) + geom_histogram()
ggplotly(p9)
#aovp(limit_bal~education, data=credit_score)
```

```{r}
subset <- credit_score %>% filter(education==2|education==1)
```

The result we received tells us that statistically, mean ages *are* different, but from the confidence interval value we can see that **this difference is hardly practically signigicant.**

Now let's see if the gender composition for the two groups differ.

```{r}
good <- filter(credit_score, default == 0)
bad <- filter(credit_score, default == 1)
c(ngoodmen, total_good, nbadmen, total_bad) %<-% c(table(good$sex)[1], sum(table(good$sex)), table(bad$sex)[1], sum(table(bad$sex)))
diffscoreci(ngoodmen, total_good, nbadmen, total_bad, conf.level = 0.95)
```

```{r}
library("ggpubr")
ggboxplot(credit_score, x = "education", y = "limit_bal", 
          color = "education", 
          order = c(6, 5, 4, 3, 2, 1),
          ylab = "Limit Balance", xlab = "Education Level")

p8 <- ggplot(credit_score, aes(x=education, y=limit_bal)) + geom_boxplot()
ggplotly(p8)
```

Let's try and test ANOVA model for comparing means of limit_bal dependent on education. ANOVA assumes homoscedasticity as well as normality of distribution. Let's build a model and view a QQ-plot, it will help us understand if the target is normally distributed.

```{r anova}
res.aov = aov(limit_bal ~ education, data=credit_score[sample(5000), ])
summary(res.aov)
TukeyHSD(res.aov)
```

```{r qq-plot not transformed}
plot(res.aov, 1)
plot(res.aov, 2)
```

As we can see, distribution is far from normal. Let's validate this hypothesis using Shapiro-Wilk test.

```{r shapiro-wilk}
aov_residuals <- res.aov[['residuals']]
shapiro.test(x=aov_residuals)
```

As expected, H_0 have to be rejected: distribution of limit_bal is not normal. That means, that we can not rely on the ANOVA model conclusions and we need to use other non-parametric criteria, for example, permutation test. Education feature is categorical (not binary) variable, which means that we have to incorporate multiple testing correction (Max-T method) in order to receive H_0 distribution for differences of means.

But first, let's try to calculate log and square root of the limit_bal, hoping that it will help to restore normality. Calculations are hidden, distribution didn't become normal anyway.

```{r}
credit_score$limit_bal_log = log(credit_score$limit_bal)
aov.log = aov(default~limit_bal_log, credit_score)
qqPlot(x=credit_score$limit_bal_log)
```

```{r}
credit_score$limit_bal_sqrt = sqrt(credit_score$limit_bal)
qqPlot(x=credit_score$limit_bal_sqrt)
```

```{r}
edu <- expand.grid(1:6, 1:6)
```

```{r}
selection <-
     subset(credit_score, education=="1"|education=="2", c(education, limit_bal))
outcome <- selection$limit_bal
treatment <- droplevels(selection$education)
```

```{r limit_bal~education permutation test}
#d <<- data.frame(tmp1)
#d <- rbind(d, tmp)
#colnames(d) <- nms
dl <- list()
foo <- function(edu) {
    p <-  c(edu['Var1'], edu['Var2'], edu['Var3'])
    selection <-
     subset(credit_score, education==p[1]|education==p[2], c(education, limit_bal))
      outcome <- selection$limit_bal
      treatment <- droplevels(selection$education)
      v <- permutation.test(treatment=subset$education, outcome=subset$limit_bal, n=1000)
      dl[p[3]] <- list(v)
    #cbind(d, )
}
```

```{r perm test continued}
d <- apply(edu, MARGIN=1, foo)
d <- t(matrix(unlist(d), ncol=nrow(edu), nrow=1000))
distrib <- apply(d, MARGIN=2, function(x) max(abs(x)))
p8 <- ggplot(as.data.frame(distrib), aes(x=distrib)) + geom_histogram()
ggplotly(p8)
```

That means that men do not return their credits **slightly more often** (3-6%) than women.

Now, let's see if the education level impacts default rate. First, calculate table which will show us the sizes of default and no-default groups for each education level, secondly, let's see how do these sizes differ from the expected ones, next calculate the value of the statistical criteria.

```{r chi2 goodness of fit test for default~education ratio}
crosstab <- table(credit_score$education, credit_score$default)
crosstab
```

```{r contigency table}
crosstab - chisq.test(crosstab)$expected
chisq.test(crosstab)
assocstats(crosstab)
```

From this we can conclude that education statistically and practically significant impacts both default and limit_bal, however, these dependencies are not linear (and not uniform as for default)

Finally, let's see if the marriage category impacts the default category.

```{r}
marriage_crosstab <- table(credit_score$marriage, credit_score$default)
marriage_crosstab
assocstats(marriage_crosstab)
```

For both variables (education and marriage) we see that **they statitically significant impact the default category**. However, the contigency coefficients (which tells us how strong the features are correlated) are relatively small.
